A Cost-based Approach for Fast Intrusion Detection

### 2017-06-05 tex/paper_v1.tex
>>-.-<<

This work aims at reduce the features used for misuse-based techniques\cite{chebrolu2005feature}, since the comparison between the whole packet and all the signatures can be redundant. And useful feature selection algorithms are proposed.
Also, the features on one specific dataset KDD Cup 1999 are analyzed by comparing the relevance between each features and information gain is employed to find the most discriminating features for each class\cite{kayacik2005selecting}.

@@

Ensemble learning is extensively used in intrusion detection\cite{zhang2008random}\cite{mukkamala2005intrusion}. The single model can be nearly all kinds of previous mentioned machine learning models.

@@

\subsubsection{Time Cost}
It is obvious that the time delay for retrieving features simply from a network packet and extracting information from all former similar connections within a time period are quite different. Different from\cite{lee2001real}, the connection based features are utilized according to\cite{lee2001real} in this paper.

@@

@@

Neural network models are at the bottom layer for the detection process, they trained for each protocol type and service respectively. Actually due to the less complexity for packets detection, neural network models perform nearly the same as traditional classifiers. The main reason of utilizing neural network models in this paper is the excellent CPU acceleration support for deep learning libraries, which will reduce the computational cost in some extent.

@@

@@

Thanks to CPU acceleration, it can be noticed that significant improvement have been made on computational cost utilizing neural network models, which is quite important in real time detection for the IDSs. Also the number of neurons in the hidden layer is reduced as small as possible, which also save the computational cost.

@@

However, there still remains many issues to be solved for IDSs, such as how to detect intrusion only after a few packages have been received, and how to accurately detect zero-day attacks. Further research will investigate to solve the mentioned issues.

>>>>

This work\cite{chebrolu2005feature} aims at reduce the features used for misuse-based techniques, since the comparison between the whole packet and all the signatures can be redundant. And useful feature selection algorithms are proposed.
Also, the features on one specific dataset KDD Cup 1999 are analyzed by comparing the relevance between each features and information gain is employed to find the most discriminating features for each class\cite{kayacik2005selecting}.

@@

Ensemble learning methods are extensively utilized in intrusion detection\cite{zhang2008random}\cite{mukkamala2005intrusion}. The single model can be nearly all kinds of previous mentioned machine learning models.

\vspace{6pt}

In the approach proposed in this paper, both the supervised learning and ensemble learning methods are employed, namely, the decision tree model and the neural networks model. Empirically, this combination demonstrates a great improvement compared with using single model for prediction.

@@

\subsubsection{Time Cost}
It is obvious that the time delay for retrieving features simply from a network packet and extracting information from all former similar connections within a time period are quite different. Different from\cite{lee2001real}, the connection based features are utilized according to\cite{lee2001real} in this paper. The exact levels of features are presented in Tab \ref{tab:feature}.

@@

\caption{\label{tab:feature}Levels of Features.}
\centering{}
\begin{tabular}{c|c|c}

\hline 
Level 1 Features & Level 2 Features & Level 3 Features\tabularnewline
\hline
Duration & Hot & Count \tabularnewline
Protocol type & Num failed logins & Serror rate \tabularnewline
Service & Logged in & Rerror rate \tabularnewline
Src bytes & Num compromised & Same srv rate \tabularnewline
Dst bytes & Root sell & Diff srv rate \tabularnewline
Flag & Su attempted & Srv count \tabularnewline
Land & Num root & Srv serror rate \tabularnewline
Wrong fragment & Num file creations & Srv rerror rate \tabularnewline
Urgent & Num shells & Srv diff host rate \tabularnewline
 & Num access files &  \tabularnewline
 & Num outbound cmds &  \tabularnewline
 & Is hot login &  \tabularnewline
 & Is guest login &  \tabularnewline
\hline

\end{tabular}
\end{table*}

@@

Neural network models are at the bottom layer for the detection process, they trained for each protocol type and service respectively. Actually due to the less complexity for packets detection, neural network models perform nearly the same as traditional classifiers. The main reason of utilizing neural network models in this paper is the excellent GPU acceleration support for deep learning libraries, which will reduce the computational cost in some extent.

@@

\subsection{Evaluation Methods}
Three evaluation metrics are employed in this paper to evaluate the performance of the intrusion detection system, namely, detection accuracy, false alarm rate and total cost:

\begin{itemize}
\item Detection accuracy is simply calculated by comparing the predicted results with the ground truth labels.
\item The false alarm rate(FAR) is defined as Eq \ref{eq:false-alarm}, in which $FP, TN$ represent false positive and true negative samples respectively.

\begin{equation}\label{eq:false-alarm}
FAR = \frac{FP}{FP+TN}
\end{equation}

\item The total cost is the sum of the computational cost and the time cost. 
\end{itemize}

@@

Thanks to GPU acceleration, it can be noticed that significant improvement have been made on computational cost utilizing neural network models, which is quite important in real time detection for the IDSs. Also the number of neurons in the hidden layer is reduced as small as possible, which also save the computational cost.

@@

However, as mentioned in Sec \ref{sec:futurework}, there still remains many issues to be solved for IDSs, such as how to detect intrusion only after a few packages have been received, and how to accurately detect zero-day attacks. Also, more attack simulations need to be organized to test the IDSs before they are actually employed in the real environment, since any crash in IDSs may leads to severe network problems.

After accomplishing the rest experiments, further research will focus on solving the mentioned issues in Sec \ref{sec:futurework}.
